Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of mixed binary observations

### 2016-06-10 Paper/paper.tex
updating more figs

@@

% #################
\begin{figure*}
\centering 
\includegraphics[width=1.8\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
\caption{We show here the ratio of systematic and statistical
measurement uncertainties for the binary chirp mass over the NSBH parameter 
space. Each panel shows the same as a function of BH mass and spin. Across
each row, we can see the effect of increasing the signal strength (i.e. SNR)
with the NS's deformability fixed. Down each column, we can see 
the effect of the increasing NS's tidal deformability at fixed SNR. We also
show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
or $100\%$ (labelled ``0'').
% 
Generally, these two sources of uncertainty are different and for BBHs, the
statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
that for BHNS binaries, its not much different until we get to high SNRs
$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
measurement uncertainty for $\mchirp$ could become dominated by the exclusion
of tidal effects in filtering templates.
}
\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
\end{figure*}
%
\begin{figure*}[!t]
\centering
\includegraphics[width=2\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_70_linear}
\caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
with the difference that the quantity in question here is the symmetric 
mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
not including the effects of NS's tidal deformation on GW emission can become the dominant
source of error for astrophysical searches with Advanced LIGO. However,
for quieter signals with $\rho\leq 30$, it will have minimal effects on the measurement
of $\eta$. We remind the reader that the SNRs here are all single detector values.
}
\label{fig:TN_EtaBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
\begin{figure*}
\centering
\includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
\caption{This figure shows the ratio of the systematic and statistical
measurement errors for BH spins, with other attributed identical to 
Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
Similar to the case of mass parameters, we find that below
$\rho\approx 30$, ignoring tidal effects in templates introduces minor systematic effects,
which remain subdominant to the statistical measurement uncertainties.
}
\label{fig:TN_BHspinBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
Past and future GW search and source characterization efforts with Advanced LIGO
have used (or plan to) BH-BH waveform templates for NS-BH mergers, ignoring
NS tidal effects on the emitted GWs. In this section we present a fully
Bayesian analysis of the effects of this assumption on the recovery of NSBH
masses and spins from their GW signals.

@@

Contours are drawn in each panel demarkating regions where we can constrain the

@@

\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1000_0_CI90_0}
\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}

@@

\caption{In these figures we highlight the regions of BH mass and spin plane 

@@

In Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we take the case of
moderately deformable neutron stars, with actual $\lambdans = 1000$ (left 
panel) and $1500$ (right panel), and ask how loud such a signal will have
to be for us to put meaningful constraints on the tidal deformability of NSs.
In the left panel, as a function BH mass and spin, 
we show the lowest SNR value $\rho_c$ in order for the signal to have enough
information that we can constrain $\dlambda$ below $200\%$.
\red{%
\begin{enumerate}
 \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
 \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
 full width or half width of the posterior. Such explanation fit in nicely into
 the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
\end{enumerate}
}
\hspace{-3mm}Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
(i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
results, which suggest that there is a fairly wide range of BH spins, less
wide (but non-trivial) range of BH masses, within
which if we get a moderately loud NSBH signal with $\rho\approx 23$, we constrain
$\lambdans$ within a factor of $2$ interval (i.e. within $\pm 100\%$) around
the true value. While this is not very precise, it can possibly allow us to
rule out candidates for the equation of state of neutron star matter.
% 
If we want to constrain $\lambdans$ better with a single observation, however,
we need SNRs of $30$ or higher. This can be seen from the right
panel of Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}, which is similar
to the left but for a higher $\lambdans (=1500)$ system, with a more stringent
constraint requirement that $\lambdans$ be measured within a factor of $1$ of 
its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$). Even for
the most conducive cases, we {\it need} the SNR to be above $30$ to reach our 
accuracy threshold (which is more than $3\times$ less likely to happen than
getting a signal with $\rho=20$).

Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
At $\rho=20$, the left panel shows that we need high BH spins 
$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
NS with $\lambdans=200$ one could even put a $\pm 50\%$ error-bar on 
$\lambdans$ measurement.

>>>>

\red{SOMEHOW INTRODUCE SNR AND MENTION IF IT IS A NETWORK SNR OR WHAT KIND OF 
'SNR' IS IT?}

@@

% #################
\begin{figure*}
\centering 
\includegraphics[width=1.6\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
\caption{We show here the ratio of systematic and statistical
measurement uncertainties for the binary chirp mass over the NSBH parameter 
space. Each panel shows the same as a function of BH mass and spin. Across
each row, we can see the effect of increasing the signal strength (i.e. SNR)
with the NS's deformability fixed. Down each column, we can see 
the effect of the increasing NS's tidal deformability at fixed SNR. We also
show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
or $100\%$ (labelled ``0'').
% 
Generally, these two sources of uncertainty are different and for BBHs, the
statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
that for BHNS binaries, its not much different until we get to high SNRs
$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
measurement uncertainty for $\mchirp$ could become dominated by the exclusion
of tidal effects in filtering templates.
}
\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
\end{figure*}
%
\begin{figure*}[!t]
\centering
\includegraphics[width=2\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_70_linear}
\caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
with the difference that the quantity in question here is the symmetric 
mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
not including the effects of NS's tidal deformation on GW emission can become the dominant
source of error for astrophysical searches with Advanced LIGO. However,
for quieter signals with $\rho\leq 30$, it will have minimal effects on the measurement
of $\eta$. We remind the reader that the SNRs here are all single detector values.
}
\label{fig:TN_EtaBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
\begin{figure*}
\centering
\includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
\caption{This figure shows the ratio of the systematic and statistical
measurement errors for BH spins, with other attributed identical to 
Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
Similar to the case of mass parameters, we find that below
$\rho\approx 30$, ignoring tidal effects in templates introduces minor systematic effects,
which remain subdominant to the statistical measurement uncertainties.
}
\label{fig:TN_BHspinBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
Past and future GW search and source characterization efforts with Advanced LIGO
have used (or plan to) BH-BH waveform templates for NS-BH mergers, ignoring
NS tidal effects on the emitted GWs. In this section we present a fully
Bayesian analysis of the effects of this assumption on the recovery of NSBH
masses and spins from their GW signals.

@@

Contours are drawn in each panel demarcating regions where we can constrain the

@@

\includegraphics[trim={0 0 0 2cm}, width=0.95\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}
\includegraphics[width=0.95\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}

@@

\caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
In these figures we highlight the regions of BH mass and spin plane 

@@

In Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we take the case of
moderately deformable neutron stars, with actual $\lambdans = 1000$ (left 
panel) and $1500$ (right panel), and ask how loud such a signal will have
to be for us to put meaningful constraints on the tidal deformability of NSs.
In the left panel, as a function BH mass and spin, 
we show the lowest SNR value $\rho_c$ in order for the signal to have enough
information that we can constrain $\dlambda$ below $200\%$.
\red{%
\begin{enumerate}
 \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
 \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
 full width or half width of the posterior. Such explanation fit in nicely into
 the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
\end{enumerate}
}
\hspace{-3mm}Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
(i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
results, which suggest that there is a fairly wide range of BH spins, less
wide (but non-trivial) range of BH masses, within
which if we get a moderately loud NSBH signal with $\rho\approx 23$, we constrain
$\lambdans$ within a factor of $2$ interval (i.e. within $\pm 100\%$) around
the true value. While this is not very precise, it can possibly allow us to
rule out candidates for the equation of state of neutron star matter.
% 
If we want to constrain $\lambdans$ better with a single observation, however,
we need SNRs of $30$ or higher. This can be seen from the right
panel of Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}, which is similar
to the left but for a higher $\lambdans (=1500)$ system, with a more stringent
constraint requirement that $\lambdans$ be measured within a factor of $1$ of 
its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$).
\red{Dont use two different thresholds in figure.} Even for
the most conducive cases, we {\it need} the SNR to be above $30$ to reach our 
accuracy threshold (which is more than $3\times$ less likely to happen than
getting a signal with $\rho=20$).

\red{%
(i) Rework text, so that you slowly build up the information needed. First, explain
the general idea pursued, then discuss one panel first, then introduce second
panel, finally, contrast both panels.
(ii) High information density. Distinction between left and right panels notation
not needed in the first sentence. deferring this info by a few sentences will make
it easier for the reader to follow the discussion.
}
\prayush{%
Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
At $\rho=20$, the left panel shows that we need high BH spins 
$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
NS with $\lambdans=2000$ one could even put a $\pm 50\%$ error-bar on 
$\lambdans$ measurement.
}
